A Method for the spectroscopic inference of fundamental stellar parameters

### 2014-05-06 tex/paper.tex
trying to get figures to work on CF
changing figs
file format
more figs
added yahapj.bst
added table back in
this is a test
merged from upstream

%\documentclass[preprint]{aastex} %double-column, single-spaced document:
\documentclass[iop,floatfix]{emulateapj}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 
\usepackage[utf8]{inputenc}

@@

\newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}

@@

@@

\section{Method}
Typically, the stellar properties we most desire are effective temperature
 $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz,
 more being developed for GAIA) and are parameterized by these fundamental
 stellar parameters, which together, we call $\vg$. 
Synthetic libraries provide high-resolution spectra which span a range of $\vg$
 covering the main sequence of spectral types. The library is typically
 specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.

In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vstar = \{\vg, \vpp\}$.

Using a synthetic stellar library as a backend and applying post-processing
 techniques, we attempt to forward-model the observed stellar spectra in order
 to determine the best-fit $\vstar$. 
Because properly sampled spectra have correlated pixels, we present a framework
 which self-consistently models these correlations.

\begin{table}[!htb]
\begin{tabular}{ll}
\hline
\hline
Symbol & Description\\
\hline
\hline
$i$ & index specifying a pixel\\
$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
  & that parameterize a synthetic spectrum from the grid\\
$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
  & are applied during ``post processing'' of the synthetic spectrum\\
$\vstar$ & $\{\vg,\vpp \}$\\
$\finst(\lambda)$ & data spectrum\\
$\fsynth(\lambda)$ & synthetic spectrum\\
$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
$\sigma_i$ & Poisson noise for a given pixel $i$\\
$R_i$ & residuals $\fDi - \fMi$\\
$\fR$ & residual vector $\fD - \fM$\\
$C$ & covariance matrix\\
$k_{\rm global}$  & global covariance kernel\\
$k_{\rm region}$  & regional covariance kernel\\
\hline
\end{tabular}
\caption{Nomenclature used in this document}
\label{tab:nomenclature}
\end{table}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{figs/model_data}
  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}. 
  \protect \todo{Model spectrum will be updated once I properly burn in the
   chain. Do these line styles look OK? I was going for something legible
   that might also print well in B/W}.}
\label{fig:model_data}
\end{center}
\end{figure*}

\subsection{Interpolation from synthetic spectral libraries}
Creating a new spectrum with a specific $\vg$ requires either synthesizing a
 new spectrum using radiative transfer codes or interpolating from nearby grid
 points in the library. 
Because it is computationally expensive to synthesize a spectrum at high
 resolution over a wide wavelength range, we choose to interpolate. 
Spline interpolation for spectra \citep{hus12}.
\todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}

\subsection{Post-processing synthetic spectra to match the data}
Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
The projected rotational velocity of the star, parameterized by $v \sin i$,
 broadens spectral line profiles and is mathematically described by convolution
 with a kernel $g_{v \sin i}(\lambda)$.  
UV, optical, and infrared spectra are acquired using a spectrograph which
 disperses light onto a CCD, defining a specific resolution and sampling rate
 for the spectrum. 
Because the resolution of the spectrograph and spacing of the pixels are
 different from the synthetic spectral library, we must convolve the raw
 spectrum with the line spread function (LSF) of the instrument $g_{\rm
 LSF}(\lambda)$ and resample to the exact pixels of the CCD. In wavelength
 space, the $v \sin i$ and LSF operations are represented by the convolution of
 the synthetic spectrum with these two kernels
\begin{equation}
  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) \otimes \fsynth(\lambda)
  \label{eqn:convolution}
\end{equation}
Using the convolution theorem, we can rewrite these operations as
 multiplications in Fourier space 
\begin{equation}
  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, {\rm synth}}(s)
\end{equation}
 where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform
 pairs. 
In order to use the fast Fourier transform (FFT) to execute these
 operations using discrete samples of the synthetic spectrum $\fsynth$, we must
 first resample $\fsynth$ so that it is on a uniform grid. 
In the case of a spectrum, the natural uniform grid is one that is equally
 sampled in velocity space, such there is an equal velocity shift $\Delta v$
 between pixels. This results in a wavelength grid that is linearly sampled in
 the logarithm of wavelength. 
Therefore, the Fourier coordinate $s$ is the number of cycles per sampling
 interval, having units of inverse velocity $[{\rm s}/{\rm km}]$. 
Next, we multiply the Fourier transform of the synthetic spectrum by the
 Fourier transforms of the rotational velocity and line spread function kernels.
Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda,
 {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength
 locations of the pixels in the detector ($\finst$). 
When resampling the spectrum, we are careful to ensure band-limited
 interpolation using splines in order to prevent introducing non-physical high
 frequency structure into the spectrum.

Finally we apply corrections for interstellar extinction and Doppler shift.
Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which
is wavelength-dependent. 
We use the (name here) dust-extinction model, which is parameterized by
 extinction in the $V$ photometric band, $A_V$. 
We correct for a radial velocity difference $v_z$ between the star and the
 earth by Doppler shifting the model spectrum. 
\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in
 Fourier space, but it's pretty fast as-is.}

\todo{We might try implementing the convolution in wavelength space, rather
 than the FFT. 
This would allow the LSF to change with wavelength.\\} 
Pixel width effects are not important as long as the LSF of the instrument is
 adequately sampled ($\gtrsim 3$ pixels across the LSF). 
If our spectrum were not properly sampled, we would need to additionally
 convolve with a boxcar the width of the pixels.

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/badlines}
  \caption{The variety of residual behaviour, depending on the quality of model
    fit. 
    From left to right: \textbf{Class 0} covariance results from slight
     model mismatch, and correlates nearby residuals. 
    \textbf{Class I}: A missing absorption line in the model leaves a large,
     highly correlated patch of negative residuals. 
    \textbf{Class II}: An extraneous line in the model leaves a large, highly
     correlated patch of positive residual.  
    \textbf{Class III}: If lines are present in the model but of the wrong
     strength, many correlated residuals of moderate amplitude will result. 
    The difficulty with class III lines is that for any specific line, there might
     exist a $\vstar$ that will fit the line, but there does not exist a
     $\vstar$ that will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure*}

Due to small imperfections in the flux-calibration of the data spectrum, there
 may be broad regions of the spectrum which are offset from the true continuum
 level by a slight amount. 
A traditional approach to address this problem is to normalize both the data
 and the spectrum to the continuum level before comparison. 
While this may work well for hotter stars with well-defined continuua, for
 cooler stars with large molecular features, normalization may introduce
 artificial features into the spectrum when the pseudo-continuum is incorrectly
 placed. 
Instead, we choose to multiply the model spectrum by a low-order Chebyshev
 polynomial that is normalized to unity. 
The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters
 in our model that we will solve for. 
If the behavior of the spectrograph is characterized by observations of
 spectrophotometric standards, then we can put priors on the coefficients that
 will prevent overfitting of the polynomials. 
For an example of a typical spectrum, and Chebyshev polynomial, see
 Figure~\ref{fig:model_data}.

@@

\subsection{Likelihood function}
To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we
 use a standard multidimensional Gaussian likelihood function which allows for
 covariance between data points. 
If the vector of residuals is 
\begin{equation}
  \fR(\vstar) = \fD - \fM(\vstar)
\end{equation}
 with a length of $N$ data points, then the likelihood function is
\begin{equation}
  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2} \fR^T C^{-1} \fR \right ) 
\end{equation}
 and the log-likelihood function is
\begin{equation}
    \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  - \frac{N}{2} \ln 2 \pi 
    \label{eqn:lnprob}
\end{equation}
$C$ is the covariance matrix which describes the covariance between pixels in
 the spectrum. 
For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
\begin{equation}
  C = 
  \begin{bmatrix}
    \sigma_0^2 & 0  & \cdots & 0\\
    0 & \sigma_1^2 & \cdots & 0\\
    \vdots  &   & \ddots  & \vdots \\
    0 & 0 & \cdots & \sigma_N^2\\
  \end{bmatrix}
  \label{eqn:covariance_diagonal}
\end{equation}
with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the
 familiar $\chi^2$ form of a sum over the square of the residuals, weighted by
 the inverse variance of each data point
\begin{equation}
  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i \frac{R_i^2}{\sigma_i^2}
\end{equation}

@@

\subsection{Parameterizing the covariance structure}
We parameterize the covariance structure using covariance kernels, which sets
 the covariance between two pixels $\lambda_i$ and $\lambda_j$. 
This behavior is analogous to the two-point correlation function used in
 cosmology, where the distance between two galaxies is used instead of
 wavelength.

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matern_matrix}
\includegraphics[width=0.4\textwidth]{figs/matern_draw}
\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel
 and typical parameters for our dataset.  
 \textbf{Right} general spectroscopic residuals near a stellar continuum
  overlaid with a sample draw using the covariance matrix, showing that the two
  look similar in structure and amplitude. 
  \protect \todo{Prettify these figures. Consistent labelling. Change
  color stretch so Matrix zeros are white.}}
\label{fig:matern}
\end{center}
\end{figure*}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance}
\includegraphics[width=0.4\textwidth]{figs/random_draw}
\caption{\textbf{Left} a covariance matrix generated with the region kernel.
\textbf{Right} general spectroscopic residuals near a mismatched stellar line
 overlaid with a sample draw using the covariance matrix, showing that the two
 look similar in structure and amplitude. 
\protect \todo{Similar improvements to
  Figure~\ref{fig:matern}.}}
\label{fig:region}
\end{center}
\end{figure*}

@@

\bibliography{stellarspectra}
\bibliographystyle{hapj}
\end{document}

>>>>

\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

\usepackage[breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}
\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 
\usepackage[utf8]{inputenc}

@@

\newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}
\newcommand{\comm}[1]{ \textcolor{Red}{SA: #1}}

@@

\graphicspath{{figs/}}

@@

\section{Method}

\comm{This section needs a true introduction of its own.  So far, you just sort 
of fly right into the details.  The introduction should basically summarize the 
underlying approach you are going to take, and then literally lay out a basic 
roadmap for the following subsections.  Some suggested paragraphs below.}

\comm{(1) describe the {\it philosophy} behind the approach: you want to make 
inferences on fundamental stellar parameters using the vast amount of 
information available in a spectrum, but at the same time you want to 
acknowledge and pro-actively account for both intrinsic parametric degeneracies 
and any practical covariances introduced into the data and the way its 
treated.  Ideally, this can be done in a flexible, modular way to easily 
incorporate additional complexity to treat different kinds of problems.  By 
casting the problem in a specific (Bayesian) framework, you can do that while 
also simultaneously learning in a data-driven way about how the underlying 
models of stellar spectra might be improved.  [Note that this paragraph might 
seem repetitive in the paper, but it sort of encapsulates all the key points so 
it merits reinforcement.]}

\comm{(2) now give an overview of the basic ``steps" involved in the process; 
in essence this is a $\sim$1-2 sentence summary of each subsequent subsection.  
This is to give the reader a ``big picture" overview of the methodology.  If we 
exclude this, then the following subsections become very confusing and 
incoherent, even for an expert in the subject.  At the same time, you want to 
be specific enough that its clear what you are going to do: (a) for a given set 
of stellar parameters, you generate (somehow) a synthetic model spectrum 
(Section 2.1); (b) you then post-process that spectrum for realism (using 
additional parameters) and to match the data (using nuisance parameters) 
(Section 2.2); (c) you then perform a pixel-by-pixel comparison with the data 
using a prescribed likelihood function (Section 2.3) and a parametric treatment 
of the covariance matrix (Section 2.4); and then (d) you iterate this procedure 
in an MCMC simulation to explore the posterior PDF (Section 2.5).  At the end, 
you want to be specific about what you ``get" from this procedure: robust 
estimates of the stellar parameters (and associated nuisance parameters), and a 
characterization of the covariance matrix that could (in principle) be used to 
identify and quantify the key discrepancies between the data and the models.}

\subsection{Synthetic Model Spectra (\comm{note change})}

\comm{This subsection needs to make a little more effort to be generic first.  
Your initial instinct is fine: first introduce the (three) fundamental 
parameters you care about (but note that these could be expanded in principle, 
to include things like B-fields, etc.).  But then, we do not want to state a 
``preferred" way of generating a synthetic model spectrum.  Instead, it would 
be interesting to make a few comments here harking back to the different 
approaches you have outlined (or will outline) in Sect 1 -- you can (a) run a 
stellar structure + radiative transfer code to generate a spectrum 
corresponding to these parameters (brute force; impractical computationally); 
(b) interpolate from a grid of stellar structures and run a simple radiative 
transfer code (i.e., SME); or (c) interpolate from a library of synthetic 
spectra (i.e., SPC, maybe others...).  Comment on the trade-off for speed 
versus accuracy (very briefly), and then say that we are going to focus here on 
the latter approach (although fundamentally the methodology is applicable to 
all three equally).}

Typically, the stellar properties we most desire are effective temperature
 $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz,
 more being developed for GAIA) and are parameterized by these fundamental
 stellar parameters, which together, we call $\vg$. 
Synthetic libraries provide high-resolution spectra which span a range of $\vg$
 covering the main sequence of spectral types. The library is typically
 specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.
\comm{Note the re-arrangement here (below); I think this subsection should deal 
only with ``obtaining" a full-resolution synthetic model spectrum with the key 
stellar parameters.  So, I've moved stuff about $v \sin i$, etc. into Sect 2.2.}

\comm{These sentences should be recast and moved into the Sect 2 introduction.}
Using a synthetic stellar library as a backend and applying post-processing
 techniques, we attempt to forward-model the observed stellar spectra in order
 to determine the best-fit $\vstar$. 
Because properly sampled spectra have correlated pixels, we present a framework
 which self-consistently models these correlations.

\begin{table}[!htb]
\begin{tabular}{ll}
\hline
\hline
Symbol & Description\\
\hline
\hline
$i$ & index specifying a pixel\\
$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
  & that parameterize a synthetic spectrum from the grid\\
$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
  & are applied during ``post processing'' of the synthetic spectrum\\
$\vstar$ & $\{\vg,\vpp \}$\\
$\finst(\lambda)$ & data spectrum\\
$\fsynth(\lambda)$ & synthetic spectrum\\
$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
$\sigma_i$ & Poisson noise for a given pixel $i$\\
$R_i$ & residuals $\fDi - \fMi$\\
$\fR$ & residual vector $\fD - \fM$\\
$C$ & covariance matrix\\
$k_{\rm global}$  & global covariance kernel\\
$k_{\rm region}$  & regional covariance kernel\\
\hline
\end{tabular}
\caption{Nomenclature used in this document \comm{Obviously this table does not 
belong here...}}
\label{tab:nomenclature}
\end{table}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{figs/model_data.pdf}
  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}. 
  \protect \todo{Model spectrum will be updated once I properly burn in the
   chain. Do these line styles look OK? I was going for something legible
   that might also print well in B/W}.}
\label{fig:model_data}
\end{center}
\end{figure*}

Creating a new spectrum with a specific $\vg$ requires either synthesizing a
 new spectrum using radiative transfer codes or interpolating from nearby grid
 points in the library. 
Because it is computationally expensive to synthesize a spectrum at high
 resolution over a wide wavelength range, we choose to interpolate. 
Spline interpolation for spectra \citep{hus12}.
\todo{Implement band-limited interpolation from the synthetic grid using spline
 interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be
 Nyquist sampling. \comm{I think this is an important thing to do {\it now}.}}
\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for
 interpolating from a simulation grid.  \comm{I do not think this is essential, 
so it should be put off until the paper is basically done, if even then.}}

\subsection{Post-Processing}

\comm{You need to incorporate these statements (moved from previous subsection) 
somewhere into the paragraph below.}
In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vstar = \{\vg, \vpp\}$.

Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
The projected rotational velocity of the star, parameterized by $v \sin i$,
 broadens spectral line profiles and is mathematically described by convolution
 with a kernel $g_{v \sin i}(\lambda)$.  
UV, optical, and infrared spectra are acquired using a spectrograph which
 disperses light onto a CCD, defining a specific resolution and sampling rate
 for the spectrum. 
Because the resolution of the spectrograph and spacing of the pixels are
 different from the synthetic spectral library, we must convolve the raw
 spectrum with the line spread function (LSF) of the instrument $g_{\rm
 LSF}(\lambda)$ and resample to the exact pixels of the CCD. In wavelength
 space, the $v \sin i$ and LSF operations are represented by the convolution of
 the synthetic spectrum with these two kernels
\begin{equation}
  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) 
   \otimes \fsynth(\lambda)
  \label{eqn:convolution}
\end{equation}
Using the convolution theorem, we can rewrite these operations as
 multiplications in Fourier space 
\begin{equation}
  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, 
    {\rm synth}}(s)
\end{equation}
 where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform
 pairs. 
In order to use the fast Fourier transform (FFT) to execute these
 operations using discrete samples of the synthetic spectrum $\fsynth$, we must
 first resample $\fsynth$ so that it is on a uniform grid. 
In the case of a spectrum, the natural uniform grid is one that is equally
 sampled in velocity space, such there is an equal velocity shift $\Delta v$
 between pixels. This results in a wavelength grid that is linearly sampled in
 the logarithm of wavelength. 
Therefore, the Fourier coordinate $s$ is the number of cycles per sampling
 interval, having units of inverse velocity $[{\rm s}/{\rm km}]$. 
Next, we multiply the Fourier transform of the synthetic spectrum by the
 Fourier transforms of the rotational velocity and line spread function kernels.
Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda,
 {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength
 locations of the pixels in the detector ($\finst$). 
When resampling the spectrum, we are careful to ensure band-limited
 interpolation using splines in order to prevent introducing non-physical high
 frequency structure into the spectrum.

\comm{It seems odd to me to do the extinction correction here, after all the 
work on velocities.  I suppose it does not really matter, but it would seem 
more harmonious if you finished discussion of all velocity-related manipulation 
and then applied both the extinction correction and the overall luminosity (or 
solid angle) scaling after that.  So, maybe re-arrange a bit for a more logical 
flow (regardless of what the code itself is doing).}
Finally we apply corrections for interstellar extinction and Doppler shift.
Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which
is wavelength-dependent. 
We use the (name here) dust-extinction model, which is parameterized by
 extinction in the $V$ photometric band, $A_V$. 
We correct for a radial velocity difference $v_z$ between the star and the
 earth by Doppler shifting the model spectrum. 
\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in
 Fourier space, but it's pretty fast as-is.  \comm{Is there a reason that would 
be better?  If this is not a bottleneck, I would ignore this.}}

\todo{We might try implementing the convolution in wavelength space, rather
 than the FFT. 
This would allow the LSF to change with wavelength.\\} 
Pixel width effects are not important as long as the LSF of the instrument is
 adequately sampled ($\gtrsim 3$ pixels across the LSF). 
If our spectrum were not properly sampled, we would need to additionally
 convolve with a boxcar the width of the pixels.  \comm{I interpret this as a 
comment to yourself...either integrate into the text in a clear way, or remove.}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/badlines.pdf}
  \caption{The variety of residual behaviour, depending on the quality of model
    fit. 
    From left to right: \textbf{Class 0} covariance results from slight
     model mismatch, and correlates nearby residuals. 
    \textbf{Class I}: A missing absorption line in the model leaves a large,
     highly correlated patch of negative residuals. 
    \textbf{Class II}: An extraneous line in the model leaves a large, highly
     correlated patch of positive residual.  
    \textbf{Class III}: If lines are present in the model but of the wrong
     strength, many correlated residuals of moderate amplitude will result. 
    The difficulty with class III lines is that for any specific line, there might
     exist a $\vstar$ that will fit the line, but there does not exist a
     $\vstar$ that will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure*}

\comm{I feel like this (important) part of the section is rushed into...it is 
lacking some kind of introductory, pedagogical lead-in to make the reader 
recognize its power and flexibility.  You are essentially touching on the big 
issue of parameterizing your calibration.  Now, we don't want to go down that 
path in detail, but it would be worth discussing by way of a better motivation 
for using these polynomials.}
Due to small imperfections in the flux-calibration of the data spectrum, there
 may be broad regions of the spectrum which are offset from the true continuum
 level by a slight amount. 
A traditional approach to address this problem is to normalize both the data
 and the spectrum to the continuum level before comparison. 
While this may work well for hotter stars with well-defined continuua, for
 cooler stars with large molecular features, normalization may introduce
 artificial features into the spectrum when the pseudo-continuum is incorrectly
 placed. 
Instead, we choose to multiply the model spectrum by a low-order Chebyshev
 polynomial that is normalized to unity. 
The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters
 in our model that we will solve for. 
If the behavior of the spectrograph is characterized by observations of
 spectrophotometric standards, then we can put priors on the coefficients that
 will prevent overfitting of the polynomials. 
For an example of a typical spectrum, and Chebyshev polynomial, see
 Figure~\ref{fig:model_data}.  \comm{One thing that has never really been clear 
to me is how you differentiate the coefficients from the solid angle scaling 
(or {\it if} you do that.  It would be worthwhile to describe that if necessary 
especially because I currently see no mention of $L_{\ast}$ or $R_{\ast}/d$ 
anywhere in this section (and of course, it should be here somewhere).}

@@

\subsection{Likelihood Calculation}

\comm{I will reserve some comments about notation and symbols for another 
iteration, but I think you at least want to be very careful about writing the 
conditional probabilities properly in Eq (5) and (7) -- it should always be 
$p(D | \theta)$ (and note that its $\theta$, not just $\vstar$).  Ultimately, I 
would like to see a brief synposis first here as well, primarily to describe 
the nomenclature up front.  E.g., what do you mean by $D$ and $M$?  I know what 
you mean, but you cannot assume the reader does without telling them.}

To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we
 use a standard multidimensional Gaussian likelihood function which allows for
 covariance between data points. 
If the vector of residuals is 
\begin{equation}
  \fR(\vstar) = \fD - \fM(\vstar)
\end{equation}
 with a length of $N$ data points, then the likelihood function is
\begin{equation}
  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2}
   \fR^T C^{-1} \fR \right ) 
\end{equation}
 and the log-likelihood function is
\begin{equation}
  \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  
   - \frac{N}{2} \ln 2 \pi 
  \label{eqn:lnprob}
\end{equation}
$C$ is the covariance matrix which describes the covariance between pixels in
 the spectrum. 
For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
\begin{equation}
  C = 
  \begin{bmatrix}
    \sigma_0^2 & 0  & \cdots & 0\\
    0 & \sigma_1^2 & \cdots & 0\\
    \vdots  &   & \ddots  & \vdots \\
    0 & 0 & \cdots & \sigma_N^2\\
  \end{bmatrix}
  \label{eqn:covariance_diagonal}
\end{equation}
with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the
 familiar $\chi^2$ form of a sum over the square of the residuals, weighted by
 the inverse variance of each data point
\begin{equation}
  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i
   \frac{R_i^2}{\sigma_i^2}
\end{equation}
\comm{I am not sure how I feel about the dividing line between Sect 2.3 and 
2.4.  It might make more sense to end right after Eq (5), and then begin Sect 
2.4 with a discussion of $C$ and the corresponding likelihood in the sample 
(independent) case, and then progress into the adopted parameterization.  I 
think I would prefer that, so please wrap up briefly after Eq (5) and push all 
this text into a re-write of Sect 2.4.  Like we discussed, you ought to just 
write as if your explanation to Tom Loredo is correct; you can always revisit 
that if need be when you hear back from him.}

@@

\subsection{Parameterizing the covariance structure}

\comm{I am leaving this alone for now, since it obviously needs more work to be 
edit-able.  Give it a shot.}

We parameterize the covariance structure using covariance kernels, which sets
 the covariance between two pixels $\lambda_i$ and $\lambda_j$. 
This behavior is analogous to the two-point correlation function used in
 cosmology, where the distance between two galaxies is used instead of
 wavelength.

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matern_matrix.png}
\includegraphics[width=0.4\textwidth]{figs/matern_draw.png}
\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel
 and typical parameters for our dataset.  
 \textbf{Right} general spectroscopic residuals near a stellar continuum
  overlaid with a sample draw using the covariance matrix, showing that the two
  look similar in structure and amplitude. 
  \protect \todo{Prettify these figures. Consistent labelling. Change
  color stretch so Matrix zeros are white.}}
\label{fig:matern}
\end{center}
\end{figure*}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance.png}
\includegraphics[width=0.4\textwidth]{figs/random_draw.png}
\caption{\textbf{Left} a covariance matrix generated with the region kernel.
\textbf{Right} general spectroscopic residuals near a mismatched stellar line
 overlaid with a sample draw using the covariance matrix, showing that the two
 look similar in structure and amplitude. 
\protect \todo{Similar improvements to
  Figure~\ref{fig:matern}.}}
\label{fig:region}
\end{center}
\end{figure*}

@@

\bibliographystyle{yahapj}
\bibliography{stellarspectra}
\end{document}
